---
title: "two_ball"
author: "Aldrich Paras"
date: "12/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# ODD AND EVEN


```{r}
# simulate odd AND even
# First ball must be odd and second ball must be even
# $1 bet, if win then win $2, if lose then lose $1


# expected winnings per game
two_ball <- function(n){
  winnings <- c()
  
  numbers <- seq(0,36)
  even <- seq(2,36,2)
  odd <- seq(1,36,2)
  
  for(i in 1:n){
    
  ball_one <- sample(numbers, 1, replace = FALSE)
  ball_two <- sample(numbers, 1, replace = FALSE)
  
  if(ball_one %in% odd & ball_two %in% even) {
    winnings <- c(winnings, 2)
  } else {
    winnings <- c(winnings, -1)
  }
  }
  
  expected <- mean(winnings)
  
  return(expected)
}
two_ball(100000)


```
The expected winnings per game for betting on even AND odd with $1 bet, winning $2 and losing $1 is -.29536



```{r}
library(ggplot2 )
# How much do you have to increase wiwnnings to gain positive expected earnings

pos_earning <- function(max_earnings, n){
  
  max_earnings <- seq(.5, max_earnings,.5)
  expected_earning <- c()
  
  numbers <- seq(0,36)
  even <- seq(2,36,2)
  odd <- seq(1,36,2)
  
  for (x in max_earnings){
    
    winnings <- c()
    
      for(i in 1:n){
      
      ball_one <- sample(numbers, 1, replace = FALSE)
      ball_two <- sample(numbers, 1, replace = FALSE)
      
      if(ball_one %in% odd & ball_two %in% even) {
        winnings <- c(winnings, x)
      } else {
        winnings <- c(winnings, -1)
      }
      }
    expected <- mean(winnings)
    expected_earning <- c(expected_earning, expected)
    
  }
  data <- data.frame(max_earnings,expected_earning)
  
  return(data)
}

x <- pos_earning(10,10000)
x

  
ggplot(x, aes(max_earnings, expected_earning)) +  geom_line() + geom_point() + ggtitle("Two-Ball Roulette: Expected Earnings as Payout Increases") + xlab("Payout (in $)") +ylab("Expected Value (in $)") + geom_hline( yintercept = 0, color = "green")
```
For two ball roulette starting at $.5 earnings per game incrementing at +$.5 until $10 with a constant loss of $1, expected winnings will be > $0 when earnings is $3.50.

```{r}
# Proportions of games you win 

prop_win <- function(n){
  
  numbers <- seq(0,36)
  even <- seq(2,36,2)
  odd <- seq(1,36,2)
  
  winnings <- c()
    
   for(i in 1:n){
      
      ball_one <- sample(numbers, 1, replace = FALSE)
      ball_two <- sample(numbers, 1, replace = FALSE)
      
      if(ball_one %in% odd & ball_two %in% even) {
        winnings <- c(winnings, 1)
      } else {
        winnings <- c(winnings, 0)
      }
      }
    
    prop <- (length(winnings[winnings == 1]))/ (length(winnings))

  
  return(prop)
}

proportion <-  prop_win(10000)
proportion
```
With two ball roulette where the first ball must be odd and the second must be even, the proportion of games won is ~ 23.92%.



```{r}
# Martingale Strategy Two ball for starting winning = $3 and loss = $1


orig_two_ball <- function(){
  winnings <- c()
  
  numbers <- seq(0,36)
  even <- seq(2,36,2)
  odd <- seq(1,36,2)
    
  ball_one <- sample(numbers, 1, replace = FALSE)
  ball_two <- sample(numbers, 1, replace = FALSE)
  
  if(ball_one %in% odd & ball_two %in% even) {
    winnings <- c(winnings, 3)
  } else {
    winnings <- c(winnings, -1)
  }
  return(winnings)
}


martingale_twoball <- function(n) {
  win_cases <- c()
  plays_cases <- c()
  
  for(i in 1:n){ 
  win_amount <- 0
  plays <- 0
  bet_amount <- 1
  
  while (win_amount < 10 && bet_amount < 100) {
      winnings = orig_two_ball() 
    if (winnings == -1) {
      win_amount = win_amount - bet_amount
      bet_amount = bet_amount * 2
    }
    else{
      win_amount = win_amount + bet_amount
      bet_amount = 1
    }
    plays = plays + 1
  }
  win_cases <- c(win_cases, win_amount)
  plays_cases <- c(plays_cases,plays)
  }
  expected_earnings <- mean(win_cases)
  expected_plays <- mean(plays_cases)
  
  return(c(expected_earnings, expected_plays))
}

martingale_twoball(100000)
```
Expected winnings per game is $-97.63 and the expected number of plays per iteration of the martingale system is ~19.18 plays.



