---
title: "two_ball"
author: "Aldrich Paras"
date: "12/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# ODD AND EVEN


```{r}
# simulate odd AND even
# First ball must be odd and second ball must be even
# $1 bet, if win then win $2, if lose then lose $1


# expected winnings per game
two_ball <- function(n){
  winnings <- c()
  
  numbers <- seq(0,36)
  even <- seq(2,36,2)
  odd <- seq(1,36,2)
  
  for(i in 1:n){
    
  ball_one <- sample(numbers, 1, replace = FALSE)
  ball_two <- sample(numbers, 1, replace = FALSE)
  
  if(ball_one %in% odd & ball_two %in% even) {
    winnings <- c(winnings, 2)
  } else {
    winnings <- c(winnings, -1)
  }
  }
  
  expected <- mean(winnings)
  
  return(expected)
}
two_ball(100000)


```
The expected winnings per game for betting on even AND odd with $1 bet, winning $2 and losing $1 is -.29536



```{r}
library(ggplot2 )
# How much do you have to increase wiwnnings to gain positive expected earnings

pos_earning <- function(max_earnings, n){
  
  max_earnings <- seq(.5, max_earnings,.5)
  expected_earning <- c()
  
  numbers <- seq(0,36)
  even <- seq(2,36,2)
  odd <- seq(1,36,2)
  
  for (x in max_earnings){
    
    winnings <- c()
    
      for(i in 1:n){
      
      ball_one <- sample(numbers, 1, replace = FALSE)
      ball_two <- sample(numbers, 1, replace = FALSE)
      
      if(ball_one %in% odd & ball_two %in% even) {
        winnings <- c(winnings, x)
      } else {
        winnings <- c(winnings, -1)
      }
      }
    expected <- mean(winnings)
    expected_earning <- c(expected_earning, expected)
    
  }
  data <- data.frame(max_earnings,expected_earning)
  
  return(data)
}

x <- pos_earning(10,10000)
x

  
ggplot(x, aes(max_earnings, expected_earning)) +  geom_line() + geom_point() + ggtitle("Two-Ball Roulette: Expected Earnings as Payout Increases") + xlab("Payout (in $)") +ylab("Expected Value (in $)") + geom_hline( yintercept = 0, color = "green")
```
For two ball roulette starting at $.5 earnings per game incrementing at +$.5 until $10 with a constant loss of $1, expected winnings will be > $0 when earnings is $3.50.

```{r}
# Proportions of games you win 

prop_win <- function(n){
  
  numbers <- seq(0,36)
  even <- seq(2,36,2)
  odd <- seq(1,36,2)
  
  winnings <- c()
    
   for(i in 1:n){
      
      ball_one <- sample(numbers, 1, replace = FALSE)
      ball_two <- sample(numbers, 1, replace = FALSE)
      
      if(ball_one %in% odd & ball_two %in% even) {
        winnings <- c(winnings, 1)
      } else {
        winnings <- c(winnings, 0)
      }
      }
    
    prop <- (length(winnings[winnings == 1]))/ (length(winnings))

  
  return(prop)
}

proportion <-  prop_win(10000)
proportion
```
With two ball roulette where the first ball must be odd and the second must be even, the proportion of games won is ~ 23.92%.



```{r}
# Martingale Strategy Two ball for starting winning = $3 and loss = $1


orig_two_ball <- function() {
  winnings <- c()
  
  numbers <- seq(0, 36)
  even <- seq(2, 36, 2)
  odd <- seq(1, 36, 2)
  
  ball_one <- sample(numbers, 1, replace = FALSE)
  ball_two <- sample(numbers, 1, replace = FALSE)
  
  if (ball_one %in% odd & ball_two %in% even) {
    winnings <- c(winnings, 3)
  } else {
    winnings <- c(winnings,-1)
  }
  return(winnings)
}


martingale_twoball <- function(n) {
  win_cases <- c()
  plays_cases <- c()
  
  for (i in 1:n) {
    win_amount <- 0
    plays <- 0
    bet_amount <- 1
    
    while (win_amount < 10 && bet_amount < 100) {
      winnings = orig_two_ball()
      if (winnings == -1) {
        win_amount = win_amount - bet_amount
        bet_amount = bet_amount * 2
      }
      else{
        win_amount = win_amount + bet_amount
        bet_amount = 1
      }
      plays = plays + 1
    }
    win_cases <- c(win_cases, win_amount)
    plays_cases <- c(plays_cases, plays)
  }
  expected_earnings <- mean(win_cases)
  expected_plays <- mean(plays_cases)
  
  return(c(expected_earnings, expected_plays))
}

martingale_twoball(100000)
```
Expected winnings per game is $-97.63 and the expected number of plays per iteration of the martingale system is ~19.18 plays.

```{r}
# expected winnings and plays as initial earnings per game raises

orig_two_ball_earning <- function(earn) {
  winnings <- c()
  
  numbers <- seq(0, 36)
  even <- seq(2, 36, 2)
  odd <- seq(1, 36, 2)
  
  ball_one <- sample(numbers, 1, replace = FALSE)
  ball_two <- sample(numbers, 1, replace = FALSE)
  
  if (ball_one %in% odd & ball_two %in% even) {
    winnings <- c(winnings, earn)
  } else {
    winnings <- c(winnings,-1)
  }
  return(winnings)
}


martingale_twoball_earning <- function(max_initial_earning, n) {
  winnings_distribution <- c()
  plays_distribution <- c()
  
  for (x in 1:max_initial_earning) {
    win_cases <- c()
    plays_cases <- c()
    
    for (i in 1:n) {
      win_amount <- 0
      plays <- 0
      bet_amount <- 1
      
      while (win_amount < 10 && bet_amount < 100) {
        winnings = orig_two_ball_earning(x)
        if (winnings == -1) {
          win_amount = win_amount - bet_amount
          bet_amount = bet_amount * 2
        }
        else{
          win_amount = win_amount + bet_amount
          bet_amount = 1
        }
        plays = plays + 1
      }
      win_cases <- c(win_cases, win_amount)
      plays_cases <- c(plays_cases, plays)
    } # end i
    winnings_distribution <- c(winnings_distribution, mean(win_cases))
    plays_distribution <- c(plays_distribution, mean(plays_cases))
    
  } # end max initial earning
  earnings <- seq(1:max_initial_earning)
  data <-
    data.frame(earnings, winnings_distribution, plays_distribution)
  return(data)
}

x <- martingale_twoball_earning(100, 10000)

plot(
  x$earnings,
  x$winnings_distribution,
  main = "Two-ball Martingale: Expected Earnings Distribution as initial Earnings Raises",
  type = "l",
  ylab = "Expected Earnings",
  xlab = "Initial Earnings"
)

plot(
  x$earnings,
  x$plays_distribution,
  main = "Two-ball Martingale: Expected # of plays as initial earnings raise",
  type = "l",
  ylab = "Number of Plays",
  xlab = "Initial Earnings"
)

```


```{r}
# Two-Ball labouchere system for odd and even

# starting with initial list of 1,2,3,4
# expected wins 


labouchere <- function(n) {
  cases_wins <- c()
  
  for (i in 1:n) {
    list_game = c(1, 2, 3, 4)
    winnings <- 0
    bet_amount = list_game[1] + list_game[length(list_game)]
    number_bets <- 0
    
    
    
    while (length(list_game) >= 1 && bet_amount < 100) {
      # odd and even play
      numbers <- seq(0, 36)
      even <- seq(2, 36, 2)
      odd <- seq(1, 36, 2)
      
      ball_one <- sample(numbers, 1, replace = FALSE)
      ball_two <- sample(numbers, 1, replace = FALSE)
      
      if (ball_one %in% odd & ball_two %in% even) {
        current_winnings <- bet_amount
      } else {
        current_winnings <- -bet_amount
      }
      
      
      if (current_winnings > 0) {
        winnings = winnings + current_winnings
        #remove one from each side of list
        list_game = list_game[-c(1, length(list_game))]
        
        if (length(list_game) == 1) {
          bet_amount <- list_game[1]
          
        } else {
          bet_amount <- list_game[1] + list_game[length(list_game)]
        }
      } else {
        winnings = winnings + current_winnings
        list_game = c(list_game, abs(current_winnings))
        bet_amount = list_game[1] + list_game[length(list_game)]
      }
      
      number_bets = number_bets + 1
      
    }
    
    cases_wins <- c(cases_wins, winnings)
    
  } # end for loop
  
  expected_earnings <- mean(cases_wins)
  
  return (expected_earnings)
  
}

labouchere(10000)

```

expected earnings for labouchere with two balls odd and even is  -494.0207. This is expected because the chance to win the two ball is so much lower than regular one ball odd or even.

```{r}

# proportion of wins for Labouchere System

labouchere_prop <- function(n) {
  cases_wins <- c()
  
  for (i in 1:n) {
    list_game = c(1, 2, 3, 4)
    winnings <- 0
    bet_amount = list_game[1] + list_game[length(list_game)]
    
    
    while (length(list_game) >= 1 && bet_amount < 100) {
      # odd and even play
      numbers <- seq(0, 36)
      even <- seq(2, 36, 2)
      odd <- seq(1, 36, 2)
      
      ball_one <- sample(numbers, 1, replace = FALSE)
      ball_two <- sample(numbers, 1, replace = FALSE)
      
      if (ball_one %in% odd && ball_two %in% even) {
        current_winnings <- bet_amount
      } else {
        current_winnings <- -bet_amount
      }
      
      
      if (current_winnings > 0) {
        winnings = winnings + current_winnings
        #remove one from each side of list
        list_game = list_game[-c(1, length(list_game))]
        
        if (length(list_game) == 1) {
          bet_amount <- list_game[1]
          
        } else {
          bet_amount <- list_game[1] + list_game[length(list_game)]
        }
      } else {
        winnings = winnings + current_winnings
        list_game = c(list_game, abs(current_winnings))
        bet_amount = list_game[1] + list_game[length(list_game)]
      }
      
    }
    
    cases_wins <- c(cases_wins, winnings)
    
  } # end for loop
  
  expected_earnings <- mean(cases_wins)
  prop <- length(cases_wins[cases_wins == 10]) / length(cases_wins)
  
  return (prop)
  
}

labouchere_prop(10000)



```
proportion of wins for labouchere two ball is .2359. Also expected to be lower since it is harder to convert on winning a two ball bet.


```{r}
# expected winnings by changing list of numbers ( e.g. - play games w initial list of (1,2,3), then (2,3,4) and so on. (all the way up to (34,35,36)).


labouchere_list <- function(n, initial_list_max) {
  # create initial lists
  
  expected_earnings <- c()
  plays_distribution <- c()
  
  for (x in 1:initial_list_max) {
    cases_wins <- c()
    cases_plays <- c()
    
    for (i in 1:n) {
      list_game = c(x, x + 1, x + 2, x + 3)
      winnings <- 0
      bet_amount = list_game[1] + list_game[length(list_game)]
      number_plays <- 0
      
      
      
      while (length(list_game) >= 1 && bet_amount < 100) {
        # odd and even play
        numbers <- seq(0, 36)
        even <- seq(2, 36, 2)
        odd <- seq(1, 36, 2)
        
        ball_one <- sample(numbers, 1, replace = FALSE)
        ball_two <- sample(numbers, 1, replace = FALSE)
        
        if (ball_one %in% odd & ball_two %in% even) {
          current_winnings <- bet_amount
        } else {
          current_winnings <- -bet_amount
        }
        
        
        if (current_winnings > 0) {
          winnings = winnings + current_winnings
          #remove one from each side of list
          list_game = list_game[-c(1, length(list_game))]
          
          if (length(list_game) == 1) {
            bet_amount <- list_game[1]
            
          } else {
            bet_amount <- list_game[1] + list_game[length(list_game)]
          }
        } else {
          winnings = winnings + current_winnings
          list_game = c(list_game, abs(current_winnings))
          bet_amount = list_game[1] + list_game[length(list_game)]
        }
        
        number_plays = number_plays + 1
        
      }
      
      cases_wins <- c(cases_wins, winnings)
      cases_plays <- c(cases_plays, number_plays)
    } # end for loop simulation
    
    expected_e <- mean(cases_wins)
    expected_p <- mean(cases_plays)
    
    expected_earnings <- c(expected_earnings, expected_e)
    plays_distribution <- c(plays_distribution , expected_p)
    
  } # end list index
  
  df <- data.frame(expected_earnings, plays_distribution)
  
  return (df)
  
}

x <- labouchere_list(10000, 10)


plot(
  seq(1:10),
  x$expected_earnings,
  type = "l",
  ylab = "Expected Earnings",
  xlab = "Initial List (1 = c(1,2,3,4))",
  main = "Two-ball Labouchere System Expected Earnings vs Initial List"
)

plot(
  seq(1:10),
  x$plays_distribution,
  type = "l",
  ylab = "Number of Plays",
  xlab = "Initial List (1 = c(1,2,3,4))",
  main = "Two-ball Labouchere System PLays Distribution vs Initial List"
)

```

calculate expected earnings per initial list( 1 = (1,2,3,4), 2 = (2,3,4,5), 3 = (3,4,5,6) and so on)







